---
title: "Week3 Assignment"
author: "A.S"
date: "September 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r set-options, echo=FALSE, cache=FALSE}
options(width = 3000)
library(stringr)
```



#### Question 3: Copy the introductory example. The vector name stores the extracted names.

#### 3(a). Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

```{r}
raw.data = "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"


names = unlist(str_extract_all(raw.data, "[[A-z]., ]{2,}"))
print(names)

# Get the first names
firstNames = unlist(str_extract_all(names, "[.,] [[A-z]]{2,}$|[[A-z]]{2,} "))
firstNames = unlist(str_extract_all(firstNames, "[[A-z]]{2,}"))
print (firstNames)

# Get the last names
lastNames = unlist(str_extract_all(names, "[^[.,]] [[A-z]]{2,}$|[[A-z]]{2,}, "))
lastNames = unlist(str_extract_all(lastNames, "[[A-z]]{2,}"))
print(lastNames)

# Construct a data frame and display the frame
data.frame("Fist Name" = firstNames, "Last Name" = lastNames)

# Get the titles
titles = unlist(str_extract_all(names, "[[A-z]]{2,}\\."))
print(titles)

```

####  3(b). Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
```{r}
data.frame("Name" = names, "Title Exists" = str_detect(names, titles))
```



####  3(c). Construct a logical vector indicating whether a character has a second name.
```{r}
# Get names with 2 parts
secondNames = unlist(str_extract_all(names, " [[A-z]]{1}\\.? [[A-z]]{1,}\\.?"))
secondNames = unlist(str_extract_all(secondNames,"[[A-z]]{1}\\.? [[A-z]]{1,}\\.?"))
print(secondNames)
data.frame("Name" = names, "Second Name Exists" = str_detect(names, secondNames))
```


####  4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.
#### 4(a) [0-9]+\\$

Answer: At least one digit betwen 0 and 9 followed by a dollar sign
```{r}
string4a = c("", "01\\", "0012$", "02020", "738372", "34384782347832", "34")
str_detect(string4a, "[0-9]+\\$")
```

#### 4(b) \\b[a-z]{1,4}\\b

Answer: Between 1 to 4 occurrences of any lower case letter a thru z that performs whole word only search
```{r}
string4b = c("a", "bc", "def", "ghij", "klmno", "02020", "34384782347832", "34")
str_detect(string4b, "\\b[a-z]{1,4}\\b")
```

#### 4(c) .*?\\.txt$

Answer: Any optional pattern that starts with any character including space and ends with .txt.
```{r}
string4c = c("a", "bc", "def", "ghij.", "klmn.txt", "*.txt", "pqr874238743 .txt", ".txt")
str_detect(string4c, ".*?\\.txt$")
```

#### 4(d) \\d{2}/\\d{2}/\\d{4}
Answer: Any 2 digits followed by a forward slash then any 2 digits followed by a forward slash and then any 4 digits number. This pattern can be used to check dates but it doesn't verify dates.
```{r}
string4d = c("1/1/2016", "02/02/2016", "20/20/0000", "1/1/16", "1/13/293393", "34384782347832", "34")
str_detect(string4d, "\\d{2}/\\d{2}/\\d{4}")
```

#### 4(e) <(.+?)>.+?</\\1>
Answer: Any pattern that matches a html/xml markup with open and closing tags.
```{r}
string4e = c("<html>Hello World</html>", "<html>Hello World<html>", "34384782347832", "34")
str_detect(string4e, "<(.+?)>.+?</\\1>")
```



#### Question (9) The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr


```{r}
code = "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
print(code)


# Apply each of the following character classes in R Regular expressions given in the text book
# to see which one extracts some meaninful clue from the secret message
# [:digit:], [:lower:], [:upper:], [:alpha:], [:alnum:], 
# [:punct:], [:graph:], [:blank:], [:space:], [:print:]
# It's clear that [:upper:] regular expression extracts meaningful text

x = unlist(str_extract_all(code, "[[:upper:].! ]"))
print(x)

# Concatenate all vectors into a string of characters 
y = paste(x, collapse="")
print(y)

# Replace dots (.) with spaces to obtain the secret message
z = str_replace_all(y, "[\\.]", " ")
print(z)
```
