---
title: "Week7 Assignment"
author: "A.S"
date: "October 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 3000)
```

## Week7 Assignment
##### <span style="color:red"><b> (1) Write R code, using your packages of choice, to load the information from each of the three sources (HTML, XML and JSON) into separate R data frames. Are the three data frames identical?</b></span>


```{r}
library(XML)
library(RCurl)
library(jsonlite)
```

##### <span style="color:red"><b> (2) Read the information from HTML, XML and JSON files into R.</b></span>
```{r}

```
##### <span style="color:blue"><b> (2.a) Load HTML file into a data frame</b></span>
```{r}
htmlFileURL = getURL("https://raw.githubusercontent.com/asajjad1234/DATA-607/master/books.html")
htmlFile_df = data.frame(readHTMLTable(htmlFileURL, header=TRUE))
class(htmlFile_df)
print (htmlFile_df)
```
##### <span style="color:blue"><b> (2.b) Link to HTML file</b></span>
 <a href="https://raw.githubusercontent.com/asajjad1234/DATA-607/master/books.html"><b>HTML File</b></a>

##### <span style="color:blue"><b> (2.c) Load XML file into a data frame</b></span>
```{r}
xmlFileURL = getURL("https://raw.githubusercontent.com/asajjad1234/DATA-607/master/books.xml")
xmlFile = xmlParse(xmlFileURL)
xmlFile_df = data.frame(xmlToList(xmlFile))
class(xmlFile_df)
print(xmlFile_df)


```

##### <span style="color:blue"><b> (2.d) Link to XML file</b></span>
 <a href="https://raw.githubusercontent.com/asajjad1234/DATA-607/master/books.xml"><b>XML File</b></a>

##### <span style="color:blue"><b> (2.e) Load JSON file into a data frame</b></span>
```{r}
jsonFileURL = getURL("https://raw.githubusercontent.com/asajjad1234/DATA-607/master/books.json")
jsonData_df = data.frame(fromJSON(jsonFileURL, flatten = FALSE))
class(jsonData_df)
print(jsonData_df)
```

##### <span style="color:blue"><b> (2.f) Link to JSON file</b></span>
 <a href="https://raw.githubusercontent.com/asajjad1234/DATA-607/master/books.json"><b>JSON File</b></a>

##### <span style="color:red"><b> (3) Conclusion.</b></span>
<span style="color:blue"><b>The files contain the same contents about books in three different formats (HTML, XML and JSON). But when they are loaded as R data frames, they look different.</b></span>


