---
title: "Week2 Assignment"
author: "A.S"
date: "September 9, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 3000)
```

#### Week2 Assignment (Movie Ratings)
Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.

Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.

This is by design a very open ended assignment.  A variety of reasonable approaches are acceptable.  You can (and should) blank out your SQL password if your solution requires it; otherwise, full credit requires that your code is “reproducible,” with the assumption that I have the same database server and R software.

You may work in a small group on this assignment.   If you work in a group, each group member should indicate who they worked with, and all group members should individually submit their week 2 assignment.

Please start early, and do work that you would want to include in a “presentations portfolio” that you might share in a job interview with a potential employer!  You are encouraged to share thoughts, ask, and answer clarifying questions in the “Week 2: R and SQL” forum.


```{r}
library(RMySQL)
library(sqldf)

mydb = dbConnect(MySQL(), user='root', password='root', dbname='movie_ratings', host='localhost')
dbListTables(mydb)

#"movie_info"
#"genre_info"  
#"director_info" 
#"actor_info" 
#"movie_actor_map"
#"movie_genre_map" 



movie_info_rs = dbSendQuery(mydb, "select * from movie_info")
movie_info_df = fetch(movie_info_rs, n=-1)

genre_info_rs = dbSendQuery(mydb, "select * from genre_info")
genre_info_df = fetch(genre_info_rs, n=-1)

director_info_rs = dbSendQuery(mydb, "select * from director_info")
director_info_df = fetch(director_info_rs, n=-1)

actor_info_rs = dbSendQuery(mydb, "select * from actor_info")
actor_info_df = fetch(actor_info_rs, n=-1)

movie_actor_map_rs = dbSendQuery(mydb, "select * from movie_actor_map")
movie_actor_map_df = fetch(movie_actor_map_rs, n=-1)

movie_genre_map_rs = dbSendQuery(mydb, "select * from movie_genre_map")
movie_genre_map_df = fetch(movie_genre_map_rs, n=-1)

print(movie_info_df)
print(genre_info_df)
print(director_info_df)
print(actor_info_df)
print(movie_actor_map_df)
print(movie_genre_map_df)

```

#### Join movie_info and actor_info tables and display results
#### Find all the actors in movie 'War Dogs'

```{r}
movie_actor_map_df = sqldf("select * from movie_info_df a, actor_info_df b, movie_actor_map_df c 
                    where a.id = c.movie_id and b.id = c.actor_id and a.title = 'War Dogs'
                    ")

print(movie_actor_map_df)
```

#### Find all movies with Ratings 4
```{r}

movie_ratings_df = sqldf("select * from movie_info_df a where a.ratings = 4")
print(movie_ratings_df)

```

#### Find all in genre Action
```{r}
movie_genre_map_df = sqldf("select * from movie_info_df a, genre_info_df b, movie_genre_map_df c 
                    where a.id = c.movie_id and b.id = c.genre_id and b.name = 'Action'
                    ")

print(movie_genre_map_df)
```


