---
title: "Week10 Assignment"
author: "A.S"
date: "November 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 3000)
```


## <span style="color:blue"><b>Document Classification (Text Mining)</b></span>

##### <span style="color:red"><b>The task of this assignment is to classify documents using already classified training documents. For this purpose documents from this site https://spamassassin.apache.org/publiccorpus/ are used.</b></span>

##### <span style="color:red"><b>The objective of this homework assignment is to classify documents</b></span>

##### <span style="color:red"><b> (1) Load Libraries :</b></span>

```{r}
library(tm)
library(stringr)
library(RTextTools)
library(SnowballC)
library(caret)
library(wordcloud)

```

##### <span style="color:red"><b> (2) Download documents and Load them into Corpus :</b></span>
```{r}


# Set Options
options(stringsAsFactors = FALSE)
```


##### <span style="color:blue"><b>(2.a) Create Corpus function : </b></span>
```{r}
# create corpus
createCorpus = function(directoryName) {
  directory = paste0(getwd(), "/", directoryName)
  corpus = Corpus(DirSource(directory=directory, pattern="\\d+"))
  print(length(corpus))
  return(corpus)
}
```

##### <span style="color:blue"><b>(2.b) Clean Corpus function : </b></span>
```{r}
# clean corpus
cleanCorpus = function(corpus) {
  tempCorpus = corpus
  tempCorpus = tm_map(tempCorpus, removeNumbers)
  tempCorpus = tm_map(tempCorpus, str_replace_all, pattern="[[.]]", replacement=" ")
  tempCorpus = tm_map(tempCorpus, removeWords, words=stopwords("en"))
  tempCorpus = tm_map(tempCorpus, tolower)
  tempCorpus = tm_map(tempCorpus, stemDocument)
  tempCorpus = tm_map(tempCorpus, PlainTextDocument)
  print(length(tempCorpus))
  return(tempCorpus)
}
```

##### <span style="color:blue"><b>(2.c) Generate DocumentTermMatrix function : </b></span>
```{r}
# build Document Term Matrix
generateDTM = function(corpus) {
  dtm = DocumentTermMatrix(corpus)
  dtm = removeSparseTerms(dtm, 0.9)
  inspect(dtm[1:10, 1:2])
  return (dtm)
}
```

##### <span style="color:blue"><b>(2.d) Create Corpus : </b></span>
```{r}
hamCorpus = createCorpus("easy_ham")
spamCorpus = createCorpus("spam")

meta(hamCorpus,  tag="type") = "ham"
meta(spamCorpus, tag="type") = "spam"

totalCorpus = c(spamCorpus, hamCorpus)
totalCorpus = cleanCorpus(totalCorpus)

metaList = factor(unlist(meta(totalCorpus, "type")))
table(metaList)
```

##### <span style="color:blue"><b>(2.e) Generate DocumentTermMatrix : </b></span>
```{r}
dtm = generateDTM(totalCorpus)
print(hamCorpus)
print(spamCorpus)

metaListLength = length(metaList)
trainPartition = round(.75 * metaListLength)
```

##### <span style="color:red"><b> (3) Train model using SVM and Classify :</b></span>
##### <span style="color:blue"><b>(3.a) Make container and Train Model : </b></span>
```{r}
container = create_container(dtm, 
                             labels=as.numeric(metaList), 
                             trainSize=1:trainPartition, 
                             testSize=trainPartition:metaListLength, 
                             virgin=FALSE)

svmModel   = train_model(container, "SVM")
svmOut     = classify_model(container, svmModel)
svmSummary = create_analytics(container, svmOut)
```

##### <span style="color:blue"><b>(3.b) Take 100% spam and test it : </b></span>
```{r}
testCorpus = createCorpus("spam_2")
testCorpus = cleanCorpus(testCorpus)
testTdm = DocumentTermMatrix(testCorpus, list(dictionary=findFreqTerms(dtm)) )
```

##### <span style="color:red"><b> (4) Analyze Results</b></span>
##### <span style="color:blue"><b>(4.a) Word Cloud (TotalCorpus) : </b></span>
```{r}
wordcloud(totalCorpus, max.words = 100)
```

##### <span style="color:blue"><b>(4.b) Word Cloud (SpamCorpus) : </b></span>
```{r}
wordcloud(spamCorpus, max.words = 100)
```

##### <span style="color:blue"><b>(4.c) Word Cloud (hamCorpus) : </b></span>
```{r}
wordcloud(hamCorpus,  max.words = 100)
```

##### <span style="color:blue"><b>(4.d) Word Cloud (testCorpus) : </b></span>
```{r}
wordcloud(testCorpus, max.words = 100)
```

##### <span style="color:blue"><b>(4.e) Frequency Plot (totalCorpus) : </b></span>
```{r}
wordMatrix1 = as.data.frame( t(as.matrix(  dtm )) )
term_frequency1 = rowSums(wordMatrix1)
term_frequency1 = sort(term_frequency1, decreasing = TRUE)
barplot(term_frequency1[1:10], col="blue", las = 2)
```

##### <span style="color:blue"><b>(4.f) Frequency Plot (testCorpus) : </b></span>
```{r}
wordMatrix2 = as.data.frame( t(as.matrix(  testTdm )) )
term_frequency2 = rowSums(wordMatrix2)
term_frequency2 = sort(term_frequency2, decreasing = TRUE)
barplot(term_frequency2[1:10], col="blue", las = 2)
```

##### <span style="color:red"><b> (5) Conclusion :</b></span>
##### <span style="color:blue"><b>(5.a) In this project spam and ham data is used to create and classify the model using Support Vector Machines (SVM) and tested with another spam data. Text/Data Mining is an interesting field and useful information could be extracted using text/data mining techniques.
<br/>
<br/>
</b></span></b></span>

